Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-26 submit/fix.tex
still ongoing fiex

For $\vec{\gamma}^1$ we find  $\kappa^{-1}_2=0.03^{+0.04}_{-0.04}$, meaning that this vector's direction is consistent with being parallel to $a_X$
(also seen visually in Figure~\ref{plane:fig}.)
That is,   $\vec{\gamma}^1$  captures magnitude variations that occur given fixed $E^F(B-V)$.

@@

\subsection{Model}
\label{modelIImodel:sec}
Model~III is the same as Model~II except for the 
addition of a new term  that relates magnitudes with the SALT2 light-curve-shape parameter,
$x_1$, that is related to magnitudes with linear coefficients $\zeta$. 
What is $x_1$ reference?
The model reads as
\begin{equation}
p \sim \mathcal{N}(0, \sigma_p),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} +\zeta_{\hat{U}} x_1 + \phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si} +\zeta_{\hat{B}} x_1 + \phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} +\zeta_{\hat{V}} x_1+ \phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} +\zeta_{\hat{R}} x_1+ \phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+\zeta_{\hat{I}} x_1+ \phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv3:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust3:eqn}
\end{equation}
The only change from Model~II is the addition of the $\vec{\zeta} x_1$ terms in Eq.~\ref{ewsiv3:eqn}.

@@

\subsection{Results}

This section describes the evaluated posterior of the model parameters.  Before doing so, we show that the model provides
a fair representation of the data.
As a diagnostic that the model results do not give catastrophically poor color predictions and are not pulled by extreme
outlying data, shown in Figure~\ref{residual:fig} are plots of observed  versus
model-expectation colors (including  supernova-specific spectral features, light-curve shape $\gamma g$, and $\phi p$ terms) for four selected colors,
and the histogram of their differences.  The residuals of the data away from the model have a compact Gaussian-like appearance, and
show no bias with respect to observed color.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{../output_fix3_x1/residual.pdf} 
            \caption{Observed  versus
            Model~III-expectation colors (including  supernova-specific spectral features, $\gamma g$, and $\phi p$  terms)
for four selected colors. 
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

Results from our analysis are shown through contours of the posterior surface for pairs of global parameters grouped by filter, 
in Figures~\ref{global1:fig} -- \ref{global5:fig}.
The confidence regions are localized and unimodal, except for the combination $\sigma_p \vec{\phi}$ that
has the sign degeneracy already noted in \S\ref{results2:sec} and plotted in Figure~\ref{M1ev:fig}.
Within the finite number of generated links, the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $g$.
Each chain converges to one of the  $\vec{\phi}$--$p$ degenerate solutions, away from which it does not migrate.  The 
confined peak of the posterior surface, in addition to the MCMC-convergence tests described in  \S\ref{results1:sec}, provide evidence that the MCMC has not failed to converge.

@@

We find significant non-zero values for $\vec{\alpha}$, $\vec{\beta}$, and $\vec{\eta}$ indicating that
$EW_{Ca}$, $EW_{Si}$, and $\lambda_{Si}$ are indicators of broadband
colors at peak
as expected.  The non-zero values for $\vec{\zeta}$ demonstrate that light-curve shape $x_1$ is also an indicator
of broadband colors.

@@

The posterior visualized in
Figures~\ref{global1:fig} -- \ref{global5:fig} shows
 that $\vec{\zeta}$, the magnitude vector associated with $x_1$, is highly correlated with $\vec{\beta}$ the vector
 associated with  $EW_{Si}$.
The direct correlation between  $x_1$ and $EW_{Si}$ has already been established
\citep{2008A&A...492..535A, 2011A&A...529L...4C}
and exists
in our sample as is shown late in  this section.  It is not surprising that our analysis does not entirely distinguish between
their influence on supernova colors. Nevertheless, their posteriors are bounded and inconsistent with zero, so they do convey independent information.
It is for this reason  that Model~III is highlighted in this article over Model~II.

The posterior gives a significant magnitude influence of the $p$ parameter.
The minimum value for $\sigma_p$ over all Monte Carlo chains is 0.039~mag and its 
68\% credible interval is $\sigma_p=0.049^{+ 0.004}_{-0.003}$~mag. 
The median data color uncertainty of 0.010~mag (the maximum is 0.042~mag) so we conclude that
the is not a due to spurious noise.

The chromatic effect of $\vec{\phi}$ is visualized in Figure~\ref{phiratio:fig}, which
gives a different perspective of the posterior using the statistic $\phi_X/\phi_{\hat{I}}-1$.
Unlike with $\gamma$, we normalize $\phi$ to $\hat{U}$ because the uncertainty in $\phi_{\hat{B}}-\phi_{\hat{V}}$ is relatively large with
credible intervals that easily cross zero.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{../output_fix3_x1/phiratio.pdf}
   \caption{Model~III median values  and corresponding 68\% intervals for $\phi_X/\phi_{\hat{I}}$ in the 5 bands.
   A dotted line at zero is shown for reference.
   \label{phiratio:fig}}
\end{figure}

As with Model~I in \S\ref{modelI:sec}, all elements of the two $\vec{\gamma}$ parameter vectors 
describing intrinsic color
are significantly non-zero with 
none of the 20000 links of 
our Monte Carlo chains for $\vec{\gamma}$ extending to 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}).
The direction of the vector is shown in Figure~\ref{plane3:fig}.
The posterior gives a quadratic fraction of $0.9996^{+0.0002}_{-0.0003}$ of $\vec{\gamma}^0$ projected onto the plane defined by
the  \citetalias{1999PASP..111...63F}-derived
$a_X$ and $b_X$, the quadratic fraction of $\vec{\gamma}^1$ projected
onto the dust plane is
$0.959^{+0.015}_{-0.029}$.
The corresponding 68\% credible is $R^F_{\mathit{eff}}=2.40^{+0.15}_{-0.15}$.

@@

An intuitive way to consider the effect of $\vec{\gamma}^1$ is through the deviation from linearity in the color-magnitude diagram.
A matrix such as the one in  Eq.~\ref{trans_I:eqn} is used to transform this $\{g_0, g_1\}$ into an $\{A_V^F, E^F(B-V)\}$ pair
(modulo additive constants)  for each link in the Monte Carlo chains. A plot showing the expected values and 68\% credible intervals of these parameters
for our supernova
sample is shown in Figure~\ref{kk:fig}. 
For reference, a line that represents $R^V_{\mathit{eff}}=2.4$ is overplotted.  The majority of supernovae lie within a narrow range above the line
meaning they have $R^F > R^V_{\mathit{eff}}$,
while the remaining fraction fall in a broader range below the line for an $R^F < R^V_{\mathit{eff}}$.  
These findings are consistent with previous results:
using fixed dust-extinction models,
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
A per supernova determination of $R^F$ requires the specification of the magnitude and colors of supernovae that experience
no dust, which is unspecified in our model.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{../output_fix3_x1/avebv_synth.pdf}
      \caption{
      Expected values and 68\% credible intervals of effective $E^F(B-V)$ and $A_V^F$ after transformation from our model $g_0$ and $g_1$ parameters, for the supernova in our sample.
      Overplotted is a line with the slope expected for $R^F=2.4$.
   \label{kk:fig}}
\end{figure}

Each supernova is described by its parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $x_1$,
$E_{\gamma^0}({\hat{B}}-{\hat{V}})=(\gamma^0_{\hat{B}}-\gamma^0_{\hat{V}})k_0$, $E_{\gamma^1}({\hat{B}}-{\hat{V}})=(\gamma^1_{\hat{B}}-\gamma^1_{\hat{V}})k_1$, and $p\sigma_p\phi_{\hat{I}}$.
Their Pearson correlation coefficients  are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), p\sigma_p\phi_{\hat{I}} ) =\\
\begin{pmatrix}
\begin{array}{rrrrrrrr}
\Delta & -0.07^{+0.06}_{-0.06} & -0.11^{+0.08}_{-0.07} & -0.17^{+0.06}_{-0.06} & 0.19^{+0.08}_{-0.08} & 0.08^{+0.07}_{-0.06} & 0.23^{+0.08}_{-0.08} & -0.13^{+0.06}_{-0.06} \\
\ldots  & EW_{Ca} & 0.09^{+0.03}_{-0.03} & -0.26^{+0.03}_{-0.03} & 0.07^{+0.03}_{-0.03} & -0.11^{+0.05}_{-0.05} & -0.18^{+0.06}_{-0.06} & -0.00^{+0.08}_{-0.07} \\
\ldots & \ldots &EW_{Si} & -0.12^{+0.03}_{-0.03} & -0.84^{+0.01}_{-0.01} & -0.13^{+0.04}_{-0.04} & 0.27^{+0.08}_{-0.09} & -0.00^{+0.08}_{-0.08} \\
\ldots  & \ldots  & \ldots  &  \lambda_{Si} & 0.14^{+0.03}_{-0.03} & 0.01^{+0.05}_{-0.05} & 0.00^{+0.06}_{-0.06} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  &\ldots & \ldots & x_1 & 0.13^{+0.04}_{-0.04} & -0.20^{+0.09}_{-0.09} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  & \ldots  &\ldots  & \ldots  & E_{\gamma^0}({\hat{B}}-{\hat{V}})& 0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} \\
\ldots  & \ldots &\ldots  & \ldots  & \ldots & \ldots  &E_{\gamma^1}({\hat{B}}-{\hat{V}})  & 0.00^{+0.07}_{-0.07} \\
\ldots  & \ldots & \ldots & \ldots  & \ldots  & \ldots  &\ldots  & p\sigma_p\phi_{\hat{I}} \\
\end{array}
\end{pmatrix}.
\end{multline}
To aid visualization, only the upper triangular part of the otherwise symmetric matrix is shown and the corresponding feature label is given on
the diagonal.

We find $>2 \sigma$ correlations between our extrinsic color-excess parameters and the input features  $EW_{Ca}$,
$EW_{Si}$, and $x_1$.
The correlation of $x_1$ with with host-galaxy (including dust) properties is known 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
There is no correlation between the  intrinsic parameter $p$ and the other features.
As alluded to earlier, the feature $EW_{Si}$ is strongly correlated with $x_1$.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

>>>>

In the above context, the direction of the $\vec{\gamma}^1$ vector is described by $\kappa_2$.
The prior that $g_0$ and  $g_1$ are uncorrelated leads  to  $\kappa_2$ being determined
by the sample 
distribution of $A^F_V$ and $E^F(B-V)$.  In this article, we do not further consider the direction of
$\vec{\gamma}^1$ but rather focus on the distribution of the per-supernova parameters.

%
%We now consider the direction of $\vec{\gamma}^1$. 
%Different supernovae are expected to encounter dust with different $R^F$;
%having  associated $\vec{\gamma}^0$ with
%the extinction from a single effective
%$\langle R^F_{\mathit{eff}}\rangle$, $\vec{\gamma}^1$ is associated with perturbations
%away from that effective value.
%Let us introduce a new parameter for each supernova, $\delta R$, such that 
%$A^F =\langle R^F_{\mathit{eff}}\rangle(1-  \delta R) E^F(B-V)$.
%Variations in the value of  $\delta R$, with the prior of the non-correlation between 
%the coefficients in Eqn.~\ref{newdust:eqn},
%lead 
%to 
%\begin{equation}
%\kappa_2 =\langle R^F_{\mathit{eff}}\rangle \left(1- \frac{\mbox{Var}(\delta R \times E^F(B-V))}{\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))}\right).
%\end{equation}
%From the matrix elements of Eqn.~\ref{trans_I:eqn},  $\kappa_2=-5.27/0.72=-7.31$, so that $\mbox{Var}(\delta R \times E^F(B-V)) = 4.00\times\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))$.  This result indicates that there is a small but significant covariance between  extinction contributions due to
%fractional deviations from an effective $R^F$ and
%the color-excess parameter $E^F(B-V)$.
%
%
%For $\vec{\gamma}^1$ we find  $\kappa^{-1}_2=0.03^{+0.04}_{-0.04}$, meaning that this vector's direction is consistent with being parallel to $a_X$
%(also seen visually in Figure~\ref{plane:fig}.)
%That is,   $\vec{\gamma}^1$  captures magnitude variations that occur given fixed $E^F(B-V)$.

@@

\subsection{Model}
\label{modelIImodel:sec}
Model~III is the same as Model~II except for the 
addition of a new term  that relates magnitudes with the SALT2 light-curve-shape parameter,
$x_1$, that is related to magnitudes with linear coefficients $\zeta$. 
The model reads as
\begin{equation}
p \sim \mathcal{N}(0, \sigma_p),
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{Ca} + \beta_{\hat{U}} EW_{Si} + \eta_{\hat{U}} \lambda_{Si} +\zeta_{\hat{U}} x_1 + \phi_{\hat{U}} p\\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{Ca} + \beta_{\hat{B}} EW_{Si} + \eta_{\hat{B}} \lambda_{Si} +\zeta_{\hat{B}} x_1 + \phi_{\hat{B}} p\\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{Ca} + \beta_{\hat{V}} EW_{Si} + \eta_{\hat{V}} \lambda_{Si} +\zeta_{\hat{V}} x_1+ \phi_{\hat{V}}  p\\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{Ca} + \beta_{\hat{R}} EW_{Si} + \eta_{\hat{R}} \lambda_{Si} +\zeta_{\hat{R}} x_1+ \phi_{\hat{R}} p\\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{Ca} + \beta_{\hat{I}} EW_{Si}+ \eta_{\hat{I}} \lambda_{Si}+\zeta_{\hat{I}} x_1+ \phi_{\hat{I}}  p
\end{pmatrix},
\label{ewsiv3:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
{\hat{U}}_o\\{\hat{B}}_o\\ {\hat{V}}_o\\{\hat{R}}_o\\{\hat{I}}_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
{\hat{U}} +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\{\hat{B}} +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
{\hat{V}}+\gamma^0_{{\hat{V}}} g_0+\gamma^1_{{\hat{V}}} g_1\\{\hat{R}}+\gamma^0_{{\hat{R}}} g_0 + \gamma^1_{{\hat{R}}} g_1\\{\hat{I}}+\gamma^0_{{\hat{I}}} g_0+\gamma^1_{{\hat{I}}} g_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust3:eqn}
\end{equation}
The only change from Model~II is the addition of the $\vec{\zeta} x_1$ terms in Eq.~\ref{ewsiv3:eqn}.

@@

\subsection{Results}
\subsubsection{Analysis Method and Validation}

The algorithm is extensively tested with simulated data with known input signals and measurement uncertainties similar to those of our dataset.
Stacks of the posteriors from 100 simulated datasets  exhibit no biases that are significant relative to the statistical uncertainties of a single posterior.

The posterior that will be shown in  \S\ref{results3global:sec} exhibits a 
confined peak, exhibiting no evidence that the MCMC has not failed to converge.

The model provides a fair representation of the data, as seen in the difference between the data and the model prediction.
Figure~\ref{residual:fig} has plots of the observed  versus
model-expectation colors (including  supernova-specific spectral features, light-curve shape $\gamma g$, and $\phi p$ terms) for four selected colors,
and the histogram of their differences.  
There are no extreme residuals 
from catastrophically poor color predictions.  There are no apparent trends between residuals and observed
color.   The residuals of the data away from the model have a compact Gaussian-like appearance.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/residual.pdf} 
            \caption{Observed  versus
            Model~III-expectation colors (including  supernova-specific spectral features, $\gamma g$, and $\phi p$  terms)
for four selected colors. 
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

This suite of tests lend confidence in our model and the validity of the derived posterior.

\subsubsection{Global Parameter Posterior Results}
\label{results3global:sec}
Results from our analysis are shown through contours of the posterior surface for pairs of global parameters grouped by filter, 
in Figures~\ref{global1:fig} -- \ref{global5:fig}.
The confidence regions are localized and unimodal, except for the combination $\sigma_p \vec{\phi}$ that
has the sign degeneracy already noted in \S\ref{results2:sec} and plotted in Figure~\ref{M1ev:fig}.
Within the finite number of generated links, the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $g$.
Each chain converges to one of the  $\vec{\phi}$--$p$ degenerate solutions, away from which it does not migrate.

@@

\subsubsection{$\vec{\alpha}$, $\vec{\beta}$, $\vec{\eta}$,  $\vec{\zeta}$: Cofactors of Observables}
The cofactors  $\vec{\alpha}$, $\vec{\beta}$, and $\vec{\eta}$ have  significant non-zero values,  meaning that
$EW_{Ca}$, $EW_{Si}$, and $\lambda_{Si}$ are indicators of broadband
colors at peak.  The non-zero values for $\vec{\zeta}$ demonstrate that light-curve shape $x_1$ is also an indicator
of broadband colors.

@@

The posterior visualized in
Figures~\ref{global1:fig} -- \ref{global5:fig} shows
 that $\vec{\zeta}$, the magnitude vector associated with $x_1$, is highly correlated with $\vec{\beta}$ the vector
 associated with  $EW_{Si}$.
The direct correlation between  $x_1$ and $EW_{Si}$ has already been established
\citep{2008A&A...492..535A, 2011A&A...529L...4C}
and exists
in our sample as is shown later in \S\ref{results3per:sec}.  
Nevertheless the $\vec{\zeta}$ and $\vec{\beta}$   posteriors are bounded and inconsistent with zero, indicating that
 $x_1$ and $EW_{Si}$ do convey independent color information.
For this reason  Model~III is highlighted in this article over Model~II.

\subsubsection{$\vec{\gamma}^0$, $\vec{\gamma}^1$: Cofactors of Extrinsic Latent Parameters}

As with Model~I in \S\ref{modelI:sec}, all elements of the two $\vec{\gamma}$ parameter vectors 
describing extrinsic color
are significantly non-zero with 
none of the 20000 links of 
our Monte Carlo chains for $\vec{\gamma}$ extending to 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}).
The direction of the vector in the ${\hat{U}}{\hat{V}}{\hat{I}}$-subspace is shown in Figure~\ref{plane3:fig}.
The posterior gives a quadratic fraction of $0.9996^{+0.0002}_{-0.0003}$ of $\vec{\gamma}^0$ projected onto the plane defined by
the  \citetalias{1999PASP..111...63F}-derived
$a_X$ and $b_X$, the quadratic fraction of $\vec{\gamma}^1$ projected
onto the dust plane is
$0.959^{+0.015}_{-0.029}$.
The   $\vec{\gamma}$  vectors project  onto $R^F_{\mathit{eff}}=2.40^{+0.15}_{-0.15}$.
This  is consistent with the $R_V=2.5$ found in the supernova samples of \citet{2011ApJ...729...55F}, and the
center  of the range of per-supernova $R^F$ values found in the samples of \citet{2014ApJ...789...32B, 2015MNRAS.453.3300A}.
Based on these facts, the supernova ``extrinsic'' colors $\vec{\gamma}^0 g_0 + \vec{\gamma}^1 g_1$  of our model are associated dust extinction.
Further discussion on this association is given in the discussion of $g_0$ and $g_1$  in \S\ref{g0g1:sec}.

@@

\subsubsection{$\sigma_p \vec{\phi}$: Cofactor of Intrinsic Latent Parameter}

The posterior shows a significant color influence of the $p$ parameter.
The minimum value for $\sigma_p$ over all Monte Carlo chains is 0.039~mag and its 
68\% credible interval is $\sigma_p=0.049^{+ 0.004}_{-0.003}$~mag. 
This is significantly larger than the median data color uncertainty of 0.010~mag (the maximum is 0.042~mag).
The evaluation of 100 simulated null data sets with no third color parameter produces no median $\sigma_p$'s  as
high as that of the data, so we assign a $>99$\% confidence in the detection of this third color term.

The values of $\vec{\phi}$ does not change  monotonically with wavelength, as is apparent  in 
the combination $\phi_X/\phi_{\hat{V}}-1$ shown in Figure~\ref{phiratio:fig}.  They have the property that
a reddening $\hat{B}-\hat{V}$ is associated with a bluing in $\hat{V}-\hat{R}$.  The color changes are largest
in the extreme $\hat{U}$ and $\hat{I}$ bands. Absorption and re-emission of light
over broad bands are the most obvious candidate for producing this behavior.
This kind of behavior is already associated with SNe~Ia.
\citet{2006ApJ...649..939K} points out that at a temperature of 7000~K, the iron/cobalt gas in the ejecta transition
from doubly and singly ionized states, making the gas phosphorescent and efficient in redistributing energy from the UV/blue to redder
wavelengths.  Qualitatively this would result in the structure for $\phi_X/\phi_V-1$ that we measure.
We speculate that varying $^{56}$Ni mass affects the rise-time to peak $B$ brightness and hence the optical depth
of the reionization front with respect to the core of the ejecta at peak, yielding the wavelength-dependent color behavior
detected in this analysis.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=4in]{../output_fix3_x1/phiratio.pdf}
   \caption{Model~III median values  and corresponding 68\% intervals for $\phi_X/\phi_{\hat{V}}-1$ in the 5 bands.
   A dotted line at zero is shown for reference.
   \label{phiratio:fig}}
\end{figure}

We emphasize that
although the $p\sigma_p \vec{\phi}$ colors have ben referred to as ``intrinsic'' due to its placement in the model, the analysis cannot quantitatively distinguish
whether its origin is intrinsic or due to a third dust parameter or some other extrinsic effect.

\subsubsection{Per-Supernova Parameter Results}
\label{results3per:sec}
Each supernova is described by its parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $x_1$,
$E_{\gamma^0}({\hat{B}}-{\hat{V}})=(\gamma^0_{\hat{B}}-\gamma^0_{\hat{V}})g_0$, $E_{\gamma^1}({\hat{B}}-{\hat{V}})=(\gamma^1_{\hat{B}}-\gamma^1_{\hat{V}})g_1$, and $p\sigma_p\phi_{\hat{I}}$.
The distributions of the Monte Carlo links for all supernovae are shown in Figure~\ref{perobject3:fig}.
There is a core concentration in the  parameter-space, with around eight objects that occupy its outskirts.
Many outliers appear in the red tail of $E_{\gamma^0}({\hat{B}}-{\hat{V}})$, as would be expected for the (infrequent) selection of supernovae
heavily extinguished by host-galaxy dust.

The Pearson correlation coefficients of the per-supernova parameters are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, x_1, E_{\gamma^0}({\hat{B}}-{\hat{V}}), E_{\gamma^1}({\hat{B}}-{\hat{V}}), p\sigma_p\phi_{\hat{I}} ) =\\
\begin{pmatrix}
\begin{array}{rrrrrrrr}
\Delta & -0.07^{+0.06}_{-0.06} & -0.11^{+0.08}_{-0.07} & -0.17^{+0.06}_{-0.06} & 0.19^{+0.08}_{-0.08} & 0.08^{+0.07}_{-0.06} & 0.23^{+0.08}_{-0.08} & -0.13^{+0.06}_{-0.06} \\
\ldots  & EW_{Ca} & 0.09^{+0.03}_{-0.03} & -0.26^{+0.03}_{-0.03} & 0.07^{+0.03}_{-0.03} & -0.11^{+0.05}_{-0.05} & -0.18^{+0.06}_{-0.06} & -0.00^{+0.08}_{-0.07} \\
\ldots & \ldots &EW_{Si} & -0.12^{+0.03}_{-0.03} & -0.84^{+0.01}_{-0.01} & -0.13^{+0.04}_{-0.04} & 0.27^{+0.08}_{-0.09} & -0.00^{+0.08}_{-0.08} \\
\ldots  & \ldots  & \ldots  &  \lambda_{Si} & 0.14^{+0.03}_{-0.03} & 0.01^{+0.05}_{-0.05} & 0.00^{+0.06}_{-0.06} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  &\ldots & \ldots & x_1 & 0.13^{+0.04}_{-0.04} & -0.20^{+0.09}_{-0.09} & 0.00^{+0.08}_{-0.08} \\
\ldots  &\ldots  & \ldots  &\ldots  & \ldots  & E_{\gamma^0}({\hat{B}}-{\hat{V}})& 0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} \\
\ldots  & \ldots &\ldots  & \ldots  & \ldots & \ldots  &E_{\gamma^1}({\hat{B}}-{\hat{V}})  & 0.00^{+0.07}_{-0.07} \\
\ldots  & \ldots & \ldots & \ldots  & \ldots  & \ldots  &\ldots  & p\sigma_p\phi_{\hat{I}} \\
\end{array}
\end{pmatrix}.
\end{multline}
To aid readability, only the upper triangular part of the otherwise symmetric matrix is shown and the corresponding feature label is given on
the diagonal.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{../output_fix3_x1/perobject_corner.pdf} 
   \caption{
   Distributions for the supernova parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $x_1$, $E_{\gamma^0}({\hat{B}}-{\hat{V}})$,  $E_{\gamma^1}({\hat{B}}-{\hat{V}})$,  and  $p\sigma_p\phi_{\hat{I}}$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject3:fig}}
\end{figure}

An extensive discussion of the correlations between the spectral features  and light-curve parameters of the SNfactory data
set can be found in \citet{chotard:thesis, leget:thesis, 2017Chotard}. 
We here confine ourselves to noting the strong anti-correlation between $EW_{Si}$ and $x_1$, which was alluded
to in \S\ref{results3global:sec} when interpreting the correlation between the cofactors $\vec{\beta}$ and $\vec{\zeta}$ that connect them with magnitudes.

\subsubsection{$g_0$, $g_1$: Extrinsic Parameters}
\label{g0g1:sec}

There are $>2 \sigma$ correlations between our extrinsic color-excess parameters $g_0$, $g_1$
(equivalently $E_{\gamma^0}({\hat{B}}-{\hat{V}})$,  $E_{\gamma^1}({\hat{B}}-{\hat{V}})$) and the input features  $EW_{Ca}$,
$EW_{Si}$, and $x_1$.  This translates to correlations between dust and the input features, given the association
of  $\vec{\gamma} g$ with dust.
The correlation of $x_1$ with with host-galaxy (including dust) properties is known 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.

The per-supernova parameter $\gamma^0_{\hat{V}} g_0$ can be associated with the $\hat{V}$-band
extinction for an effective $R^F_{\mathit{eff}}=2.40$ dust, as
$\vec{\gamma}_0$  has nearly the same effect on color as dust extinction at that constant $R^F$.
Figure~\ref{k0_med:fig} shows the histogram of
per-supernova
median values
from the Monte Carlo chains
and the  stack of the posteriors of all supernovae  for $\gamma^0_{\hat{V}} g_0$ 
relative to that of an arbitrary supernova  $\gamma^0_{\hat{V}} g_0|_0$.
(Subtracting the value of the one supernova suppresses correlated errors in $g_0$ common to all supernovae.)
The distributions are non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with
simulations based on expected dust distributions
within galaxies and the distribution of galaxy orientations with respect to the observer,
which determine column densities along the lines of sight toward supernovae that are consistent with observations
\citep{1998ApJ...502..177H, 2007ApJ...659..122J}.  
We emphasize that unlike other analyses,
this distribution is given by the data, and so is not dependent on any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{../output_fix3_x1/deltagamma1_med.pdf}
      \caption{
      Normalized stack of the posteriors of all supernovae  and histogram 
      of  per-supernova 
median values of: (left) $\gamma^0_{\hat{V}} g_0-\gamma^0_{\hat{V}} g_0|_0$, which is associated with the $A^F_V$ of an $R^F_{\mathit{eff}} \sim 2.4$ dust model;
(right) $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\langle R^F_{\mathit{eff}}\rangle$ value.
   \label{k0_med:fig}}
\end{figure}

An instructive way to consider the  $\{g_0, g_1\}$ parameters
is through their transformation to  $\{A_V^F, E^F(B-V)\}$ (modulo additive constants)   through the
matrix in  Eq.~\ref{trans_I:eqn}.
A plot showing the expected values and 68\% credible intervals of these parameters
for our supernova
sample is shown in Figure~\ref{kk:fig}. 
For reference, a line that represents $R^V_{\mathit{eff}}=2.4$ is overplotted.  The majority of supernovae lie within a narrow range above the line
meaning they have $R^F > R^V_{\mathit{eff}}$,
while the remaining fraction fall in a broader range below the line for an $R^F < R^V_{\mathit{eff}}$.  
The negative tail in  $\gamma^1_{\hat{V}} g_1-\gamma^1_{\hat{V}} g_1|_0$ of the right plot of Figure~\ref{k0_med:fig} is directed toward smaller $R^F$.
These findings are consistent with previous results:
using fixed dust-extinction models,
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
A per supernova determination of $R^F$ requires the specification of the magnitude and colors of supernovae that experience
no dust, which is unspecified in our model.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{../output_fix3_x1/avebv_synth.pdf}
      \caption{
      Expected values and 68\% credible intervals of effective $E^F(B-V)$ and $A_V^F$ after transformation from our model $g_0$ and $g_1$ parameters, for the supernova in our sample.
      Overplotted is a line with the slope expected for $R^F=2.4$.
   \label{kk:fig}}
\end{figure}

\subsubsection{$p$: Intrinsic Parameter}
\label{p:sec}
